Trends in Preventive Care Practices among US Adults with Diabetes, 2008-2020 

### 2022-12-28 scripts/manuscript/manuscript_draft.qmd
Updated preventive table with SE, apc, and aapc

`%!in%` <- Negate(`%in%`)

@@

```{r}

preventive_practice_table <- age_adjusted_preventive %>% 
  filter(label == 'At least three preventive practices followed') %>% 
  mutate(age_adjusted_prop = age_adjusted_prop * 100) %>% 
  pivot_wider(
    id_cols = c(stratifier, variable, label, strata),
    names_from = year,
    values_from = age_adjusted_prop
  ) %>% 
  filter(strata %!in% c('Not available', 'Other Race/Not Hispanic')) %>% 
  mutate(stratifier = factor(stratifier,
                             c('overall', 'age', 'sex', 'race', 'edu', 'insurance', 'poverty')),
         strata = case_when(strata == 'stat_0' ~ '-', 
                            strata == 'Greater than HS' ~ 'Greater than high school',
                            TRUE ~ strata)) %>% 
  mutate(strata = as.factor(strata)) %>% 
  mutate(stratifier = recode_factor(
    stratifier,
    overall = 'Overall',
    age = 'Age',
    sex = 'Sex',
    race = 'Race/Ethnicity',
    edu = 'Highest degree earned',
    insurance = 'Insurance coverage',
    poverty = 'Poverty level'
  )) %>% 
  group_by(stratifier) %>%
  select(stratifier, strata, `2008`:`2020`) %>%
  gt() %>% 
  fmt_number(columns = `2008`:`2020`, decimals = 1)
preventive_practice_table$`_data` <- preventive_practice_table$`_data`[c(1:18, 20, 19, 21), ]
preventive_practice_table %>% 
  gtsave('preventive_practices.docx')
```

>>>>

`%!in%` <- Negate(`%in%`)
sigfig <- function(number, digits) {

   formatC(signif(number, digits = digits), digits = digits, format = "fg", flag = "#")

}

@@

```{r}
# Clean up the AAPC table and get it into format for printing to table
cleaned_jp_aapc <- jp_prevention_aapc %>% 
  select(-aapc_index) %>% 
  mutate(across(starts_with('aapc'),
                ~sigfig(.x, digits = 2)),
         across(starts_with('aapc'),
                ~case_when(.x == ' NA' ~ NA_character_,
                           TRUE ~ .x)), 
         AAPC = case_when(!is.na(aapc) ~ glue::glue('{aapc} ({aapc_c_i_low}, {aapc_c_i_high})'),
                          TRUE ~ NA_character_)) %>%
  select(stratifier, strata, AAPC)

# Format APC table for future joining
cleaned_jp_apc <- jp_prevention_apc %>% 
  # No model has more than a single joinpoint, so make wider and then
  # remove the excess columns to determine if there's no join point (segmend_end = 2020)
  # or a joinpoint (segment_end != 2020)
  pivot_wider(id_cols = c(stratifier, strata, model),
              names_from = segment,
              values_from = c(segment_start, segment_end, apc, p_value, apc_significant,
                              apc_95_lcl, apc_95_ucl)
  ) %>% 
  select(-segment_start_0, -segment_start_1, -segment_end_1) %>% 
  # Make segment empty if no joinpoint
  mutate(segment = case_when(segment_end_0 == 2020 ~ NA_integer_,
                             TRUE ~ segment_end_0)) %>% 
  # MAke apc empty if not significant
  mutate(across(c(apc_0, apc_95_lcl_0, apc_95_ucl_0),
                ~ case_when(p_value_0 > 0.05 & p_value_1 > 0.05 ~ NA_real_,
                            TRUE ~ .x)), 
         across(c(apc_1, apc_95_lcl_1, apc_95_ucl_1),
                ~ case_when(p_value_0 > 0.05 & p_value_1 > 0.05 ~ NA_real_,
                            TRUE ~ .x)),
         across(starts_with('apc') | starts_with('p_'),
                ~sigfig(.x, digits = 2)),
         across(starts_with('apc') | starts_with('p_'),
                ~case_when(.x == ' NA' ~ NA_character_,
                           TRUE ~ .x))) %>% 
  # Concatenate apc column with CI columns
  mutate(`APC Period 1` = case_when(
       !is.na(apc_0) ~ glue::glue('{apc_0} ({apc_95_lcl_0}, {apc_95_ucl_0})'),
       TRUE ~ NA_character_
                                    ),
    `APC Period 2` = case_when(
       !is.na(apc_1) ~ glue::glue('{apc_1} ({apc_95_lcl_1}, {apc_95_ucl_1})'),
       TRUE ~ NA_character_
    )
  ) %>%
  select(stratifier, strata, segment, starts_with('APC '))

# Clean up the age adjusted preventive practices and then join with the aapc table
clean_age_adjusted <- age_adjusted_preventive %>% 
  # Filter to just the >=3 practices
  filter(label == 'At least three preventive practices followed') %>% 
  # Filter down to every other year
  filter(year %in% seq(2000, 2020, by = 2)) %>% 
  mutate(age_adjusted_prop = age_adjusted_prop * 100,
         chr_age_adj_prop = sigfig(age_adjusted_prop, digits = 4),
         age_adjusted_se_perc = sigfig(age_adjusted_se_perc, digits = 2),
         percentage = glue::glue('{chr_age_adj_prop} ({age_adjusted_se_perc})')) %>% 
  pivot_wider(
    id_cols = c(stratifier, variable, label, strata),
    names_from = year,
    values_from = c(age_adjusted_prop, percentage),
  ) %>% 
  # Remove excess strata
  filter(strata %!in% c('Not available', 'Other Race/Not Hispanic')) %>% 
  # Relevel stratifier factor, clean up stratifier and strata names
  mutate(stratifier = factor(stratifier,
                             c('overall', 'age', 'sex', 'race', 'edu', 'insurance', 'poverty')),
         strata = case_when(strata == 'stat_0' ~ '-', 
                            strata == 'Greater than HS' ~ 'Greater than high school',
                            TRUE ~ strata),
         `Percent Change` = sigfig(((age_adjusted_prop_2020 - age_adjusted_prop_2008) / age_adjusted_prop_2008) * 100, digits = 4)) %>% 
  select(-starts_with('age_adjusted_prop')) %>% 
  setNames(., str_replace(names(.), 'percentage_', "")) %>% 
  mutate(strata = as.factor(strata)) %>% 
  mutate(stratifier = recode_factor(
    stratifier,
    overall = 'Overall',
    age = 'Age',
    sex = 'Sex',
    race = 'Race/Ethnicity',
    edu = 'Highest degree earned',
    insurance = 'Insurance coverage',
    poverty = 'Poverty level'
  )) %>% 
  # Join with AAPC table
  left_join(cleaned_jp_aapc, by = c('stratifier', 'strata'))

cleaned_preventive_df <- left_join(clean_age_adjusted, cleaned_jp_apc, by = c('stratifier', 'strata')) %>% 
  mutate(AAPC = case_when(is.na(AAPC) & is.na(`APC Period 2`) ~ `APC Period 1`,
                          TRUE ~ AAPC),
         `APC Period 1` = case_when(is.na(`APC Period 1`) & is.na(`APC Period 2`) ~ AAPC,
                                    TRUE ~ `APC Period 1`)) %>%
  # Rename poverty levels
  mutate(strata = case_when(as.character(strata) == 'Poor/Negative' ~ 'Income < 100% of the poverty line',
                            as.character(strata) == 'Low income' ~ 'Income 100% - 199% of the poverty line',
                            as.character(strata) == 'Middle income' ~ 'Income 200% - 399% of the poverty line',
                            as.character(strata) == 'High income' ~ 'Income > 400% of the poverty line',
                            TRUE ~ as.character(strata)))

preventive_practice_table <- cleaned_preventive_df %>% 
  group_by(stratifier) %>%
  select(stratifier, strata, `2008`:`2020`, `Percent Change`, `Joinpoint Year` = segment,
         `APC Period 1`, `APC Period 2`, `AAPC`) %>%
  gt() %>% 
  sub_missing(missing_text = ' ')
preventive_practice_table$`_data` <- preventive_practice_table$`_data`[c(1:18, 20, 19, 21), ]
preventive_practice_table %>% 
  gtsave('preventive_practices.docx')
```
